\documentclass[10pt,draft]{article}

\input{prelude}

\begin{document}
\title{Domain-Specific Languages for Enterprise Systems}
\author{Tom Hvitved \and Patrick Bahr \and Jesper Andersen}

% hack:
\renewcommand{\today}{\small Department of Computer Science, University of
    Copenhagen\\ Universitetsparken 1, 2100 Copenhagen, Denmark\\
  \{\texttt{hvitved,bahr,jespera}\}\texttt{@diku.dk}}

\maketitle

\begin{abstract}
  The process-oriented event-driven transaction systems (POETS)
  architecture introduced by Henglein et al. is a novel software architecture for
  enterprise resource planning (ERP) systems. POETS employs a pragmatic
  separation between
  \begin{inparaenum}[(i)]
  \item transactional data, that is what has happened;
  \item reports, that is what can be derived from the transactional
    data; and
  \item contracts, that is which transactions are expected in the
    future.
  \end{inparaenum}
  Moreover, POETS applies domain-specific languages (DSLs) for
  specifying reports and contracts, in order to enable succinct
  declarative specifications as well as rapid adaptability and
  customisation. In this report we document an implementation of a
  generalised and extended variant of the POETS architecture. The
  generalisation is manifested in a detachment from the ERP domain,
  which is rather an instantiation of the system than a built-in
  assumption. The extensions amount to a customisable data model based
  on nominal subtyping; support for run-time changes to the data
  model, reports and contracts, while retaining full
  auditability; and support for referable data that may evolve over
  time, also while retaining full auditability as well as referential
  integrity. Besides the revised architecture, we present the DSLs
  used to specify data definitions, 
  reports, and contracts respectively, and we provide the complete
  specification for a use case scenario, which demonstrates the
  conciseness and validity of our approach. Lastly, we describe
  technical aspects of our implementation, with focus on the techniques
  used to implement the tightly coupled DSLs.
\end{abstract}

\tableofcontents

\section{Introduction}
\label{sec:introduction}
\input{introduction}

\section{Revised POETS Architecture}
\label{sec:extended-poets}
\input{poets}

\section{Use Case: \texorpdfstring{\muerp}{\textmu{}ERP}}
\label{sec:use-case}
\input{usecase}

\section{Implementation Aspects}
\label{sec:implementation-aspects}
\input{implementation}

\section{Conclusion}
\label{sec:conclusion}
\input{conclusion}

\bibliographystyle{plain}
\bibliography{refs}

\clearpage

\appendix

\section{Predefined Ontology}
\label{app:predefined-ontology}
\input{ont}
\clearpage

\section{Static and Dynamic Semantics of the Report Language}
\label{sec:stat-dynam-semant}
\input{rep}
\clearpage

\section{\texorpdfstring{\muerp}{\textmu{}ERP} Specification}
\label{app:muerp}
\input{muerp}

\end{document}